FinalManuscript_DISCS_UG

### 2019-01-18 pcsc-paper/parts/01_introduction.tex
Recreated graphs with pgfplots

Applications of the Paillier cryptosystem have included such as use in linear digital image transformations \cite{ziad_cryptoimg:_2016}, and linear algebra applications \cite{hutchison_privacy-preserving_2009}. However, many extensions to the Paillier cryptosystem have also been published, allowing floating-point arithmetic. In this paper, we will summarize these extensions and concern ourselves with the approximation of the transcendental functions using the Paillier cryptosystem given input constraints shown in table \ref{tab:inputconstraints}. By demonstrating how the Paillier cryptosystem can be applied to approximate common non-linear functions, we hope to extend its usability as a tool for privacy-preserving computation.
\begin{table}
	\caption{Summary of Functions [WIP]}
	\label{tab:inputconstraints}
	\begin{tabular}{ccl}
		\toprule
		Functions & Input Constraints\\
		\midrule
		$e^x$ & $x\in[0,255]$\\
		$\log(1+x)$ & $x\in[0,255]$\\
		$\sin x$ & $x\in[-2\pi,2\pi]$ \\
		$\arctan x$ & $x\in[-255,255]$\\
	\bottomrule
\end{tabular}
\end{table}

>>>>

Applications of the Paillier cryptosystem have included such as use in linear digital image transformations~\cite{ziad_cryptoimg:_2016}, and linear algebra applications~\cite{hutchison_privacy-preserving_2009}. However, many extensions to the Paillier cryptosystem have also been published, allowing floating-point arithmetic. In this paper, we will summarize these extensions and concern ourselves with the approximation of the transcendental functions using the Paillier cryptosystem given input constraints shown in Table~\ref{tab:inputconstraints}. By demonstrating how the Paillier cryptosystem can be applied to approximate common non-linear functions, we hope to extend its usability as a tool for privacy-preserving computation.
\begin{table}
	\caption{Summary of functions [WIP]}
	\label{tab:inputconstraints}
	\begin{tabular}{ccl}
		\toprule
		Functions & Input constraints\\
		\midrule
		$e^x$ & $x\in[0,255]$\\
		$\log(1+x)$ & $x\in[0,255]$\\
		$\sin x$ & $x\in[-2\pi,2\pi]$ \\
		$\arctan x$ & $x\in[-255,255]$\\
		\bottomrule
\end{tabular}
\end{table}

### 2019-01-18 pcsc-paper/parts/02_01_paillier.tex
Recreated graphs with pgfplots

\subsection{The Paillier Cryptosystem}
The Paillier cryptosystem \cite{stern_public-key_1999}, developed by Pascal Paillier, is a probabilistic encryption scheme which is based on the composite residuosity class problem. The scheme allows for the encryption and decryption of integer messages, and is known to be additively homomorphic. We now state the encryption and decryption algorithms of the Paillier cryptosystem and its homomorphic properties.

@@

\subsubsection{Extending to Floating-Point Numbers}
We can use the following protocol described in \cite{ziad_cryptoimg:_2016}, in order to extend the privacy-preserving integer addition and integer scalar multiplication of Paillier cryptosystem to floating-point numbers.

We represent a floating-point (FP) number as a pair of two integers $(m,e)$ representing the mantissa and exponent of the FP number with respect to a base $b$. The mantissa $m$ is encrypted, while the exponent $e$ is unencrypted.
Let $a,b,c$ be FP numbers represented by the pairs $(m_a,e_a),(m_b,e_b),(m_c,e_c)$ respectively. Let $\oplus,\otimes$ represent the homomorphic operations which correspond to the addition and multiplication of integers in the Paillier cryptosystem, respectively. We define the corresponding FP number operations as follows:
\begin{description}
  \item[Addition.]
    To compute $E(c)=E(a+b)$ we compute
    \begin{align*}
      E(m_c) &= E(m_1) \oplus (b^{e_b-e_a} \otimes E(m_b)), e_c = e_a & \text{ if } e_a \leq e_b \\
      E(m_c) &= E(m_b) \oplus (b^{e_a-e_b} \otimes E(m_a)), e_c = e_b & \text{ if } e_a > e_b
    \end{align*}
  \item[Scalar multiplication.]
    To compute $E(c) = E(ab)$, where $a$ and $E(b)$ are known (i.e. $m_a$ is not encrypted), we compute
    \begin{align*}
      E(m_c) &= m_a \otimes E(m_b)\\
      e_c &= e_a + e_b
    \end{align*}
\end{description}
\subsubsection{Secure Division}
We can use the following two-party scheme defined in \cite{boukoros_lightweight_2017} to perform privacy-preserving division.
Suppose Bob has $E(x)$ and $E(y)$ and wants to obtain $E(x/y)$ without exposing the value of either integer.
\begin{itemize}
	\item Bob first selects a random number $r$ and computes $E(rx)$ and $E(ry)$. Bob can do this since $r$ is a plaintext constant.
	\item Bob sends $E(rx)$ and $E(ry)$ to Alice, who decrypts both values and computes $x/y$ in the plaintext domain.
	\item Alice encrypts and sends $E(x/y)$ to Bob.
\end{itemize}

>>>>

\subsection{The Paillier Cryptosystem}
The Paillier cryptosystem~\cite{stern_public-key_1999}, developed by Pascal Paillier, is a probabilistic encryption scheme which is based on the composite residuosity class problem. The scheme allows for the encryption and decryption of integer messages, and is known to be additively homomorphic. We now state the encryption and decryption algorithms of the Paillier cryptosystem and its homomorphic properties.

@@

\subsubsection{Extension to Floating-Point Numbers}
\label{sec:fp_operations}
We can use the following protocol described in~\cite{ziad_cryptoimg:_2016}, in order to extend the privacy-preserving integer addition and integer scalar multiplication of Paillier cryptosystem to floating-point numbers.

We represent a floating-point (FP) number as a pair of two integers $(m,e)$ representing the mantissa and exponent of the FP number with respect to a base $b$. The mantissa $m$ is encrypted, while the exponent $e$ is unencrypted.
Let $a,b,c$ be FP numbers represented by the pairs $(m_a,e_a),(m_b,e_b),(m_c,e_c)$ respectively. Let $\oplus,\otimes$ represent the homomorphic operations which correspond to the addition and multiplication of integers in the Paillier cryptosystem, respectively. We define the corresponding FP number operations as follows:
\begin{description}
  \item[Addition.]
    To compute $E(c)=E(a+b)$ we compute
    \begin{align*}
      E(m_c) &= E(m_1) \oplus (b^{e_b-e_a} \otimes E(m_b)), e_c = e_a & \text{ if } e_a \leq e_b \\
      E(m_c) &= E(m_b) \oplus (b^{e_a-e_b} \otimes E(m_a)), e_c = e_b & \text{ if } e_a > e_b
    \end{align*}
  \item[Scalar multiplication.]
    To compute $E(c) = E(ab)$, where $a$ and $E(b)$ are known (i.e. $m_a$ is not encrypted), we compute
    \begin{align*}
      E(m_c) &= m_a \otimes E(m_b)\\
      e_c &= e_a + e_b
    \end{align*}
\end{description}
\subsubsection{Secure Division}
We can use the following two-party scheme defined in~\cite{boukoros_lightweight_2017} to perform privacy-preserving division.
Suppose Bob has $E(x)$ and $E(y)$ and wants to obtain $E(x/y)$ without exposing the value of either integer.
\begin{itemize}
	\item Bob first selects a random number $r$ and computes $E(rx)$ and $E(ry)$. Bob can do this since $r$ is a plaintext constant.
	\item Bob sends $E(rx)$ and $E(ry)$ to Alice, who decrypts both values and computes $x/y$ in the plaintext domain.
	\item Alice encrypts and sends $E(x/y)$ to Bob.
\end{itemize}

### 2019-01-18 pcsc-paper/parts/03_exponentialandlogarithmic.tex
Recreated graphs with pgfplots

\section{Exponential and Logarithmic Functions}
We now present our main contribution in this paper: a general method to improve the general accuracy of the logarithm approximation given by Khattri in \cite{khattri_new_2009} for use in privacy-preserving calculation, provided that accuracy is only required for a specific range.

For this approximation, we assume we only require accurate results in the input range $x \in [0, 255]$. This reflects some potential applications such as in logarithm intensity transformations in image processing \cite{gonzalez_digital_2008}.

@@

Applying the five-point Gauss-Legendre quadrature rule to approximate $\int_{n}^{\alpha n + \alpha}{\frac{1}{x}\diff x}$ with $\alpha = 1/16$ using SageMath 8.3, we arrive at the approximation:
\begin{align}\label{eq:optimal_log_approximation}
	\begin{split}
		&\log(1+x) \\
		&=\frac{137x^5 + 26685x^4 + 617370x^3 - 6498630x^2 - 121239315x - 257804775}
		{30(x^5 + 405x^4 + 27210x^3 + 488810x^2 + 2536005x + 3122577)}\\
		&+ \log{16}
	\end{split}
\end{align}

We note that the approximation in equation \ref{eq:optimal_log_approximation} requires the value of $\log{16}$. This can be computed in the plaintext domain and added to an encrypted value using the properties of the Paillier cryptosystem. The remaining arithmetic operations in the approximation can be computed in the encrypted domain, using the protocols described in section \ref{sec:fp_operations}. Therefore, this approximation can be calculated in a privacy-preserving manner.

Figure \ref{fig:log_error_comparison} is a graph comparing the absolute error of the scaled approximation and the original approximation by Khattri given an input $x$. We can see that scaling the input by some constant factor $\alpha$ improves the accuracy of the closed-form approximation, although it requires that $\log \alpha$ be known.

\begin{figure}[!ht]
		\centering
		\includegraphics[width=.9\linewidth]{figures/log_error_comparison.png}
		\caption{Comparison of the absolute error between the original approximation in equation \ref{eq:standard_logarithm_quadrature} and the scaled approximation in equation \ref{eq:optimal_log_approximation}.}
		\label{fig:log_error_comparison}
\end{figure}

We now show how we used numerical methods to determine that a scaling factor of $\alpha = 1/16$ minimizes the maximum absolute error of the scaled approximation for $\log{(1+x)}$, within the range $x \in [0, 255]$.
Let the function $I(x,\alpha)$ be the function obtained by approximating $\int_{n}^{\alpha n + \alpha}{\frac{1}{x}\diff x}$ with five-point Gauss-Legendre quadrature. The absolute error of the approximation compared to  $\log{(1+x)}$ at a given $x$ and $\alpha$ is given by the following function.
\begin{align*}
	A(x,\alpha) = \left| I(x,\alpha) - \log\alpha - \log{(1+x)}\right|.
\end{align*}
We want to show that the function
\begin{align*}
	f(\alpha) = \max_{0\leq x \leq 255}{A(x,\alpha)}
\end{align*}
is minimized at $\alpha = 1/16$.
We consider the functions $A(0,\alpha)$ and $A(255,\alpha)$. These are graphed in figure \ref{fig:endpoint_plot}.

It can be determined numerically using SageMath 8.3 that a local minimum of $\max(A(0,\alpha),A(255,\alpha))$ on the interval $[0,1]$ is at $\alpha = 1/16$, which is found at the intersection of the two graphs.
\begin{figure}[!ht]
		\centering
		\includegraphics[width=.9\linewidth]{figures/endpoint_plot.png}
		\caption{Graphs of $A(x,\alpha)$ at $x=0$ and $x=255$.}
		\label{fig:endpoint_plot}
\end{figure}
We then consider the function $A(x,1/16)$, shown in figure \ref{fig:single_alpha_plot}.
It was determined numerically using SageMath 8.3 that the local maxima of $A(x,1/16)$ on the interval occur at $x=0$ and $x=255$. Since a scaling factor other than $alpha=1/16$ yields a higher absolute error at either $x=0$ or $x=255$, as seen in figure \ref{fig:endpoint_plot}, it follows that a scaling factor of $\alpha = 1/16$ minimizes the maximum absolute error of the scaled approximation for $\log{(1+x)}$, within the range $x \in [0, 255]$.
\begin{figure}[!ht]
		\centering
		\includegraphics[width=.9\linewidth]{figures/single_alpha_plot.png}
		\caption{Graph of $A(x,1/16)$, the absolute error of the approximation in equation \ref{eq:optimal_log_approximation} given a scaling factor of $\alpha=1/16$.}
		\label{fig:single_alpha_plot}
\end{figure}

>>>>

\section{Exponential and Logarithmic Functions}
We now present our main contribution in this paper: a general method to improve the general accuracy of the logarithm approximation given by Khattri in~\cite{khattri_new_2009} for use in privacy-preserving calculation, provided that accuracy is only required for a specific range.

For this approximation, we assume we only require accurate results in the input range $x \in [0, 255]$. This reflects some potential applications such as in logarithm intensity transformations in image processing~\cite{gonzalez_digital_2008}.

@@

Applying the five-point Gauss--Legendre quadrature rule to approximate $\int_{n}^{\alpha n + \alpha}{\frac{1}{x}\diff x}$ with $\alpha = 1/16$ using SageMath 8.3, we arrive at the approximation:
\begin{align}\label{eq:optimal_log_approximation}
	\begin{split}
		&\log(1+x) \\
		&=\frac{137x^5 + 26685x^4 + 617370x^3 - 6498630x^2 - 121239315x - 257804775}
		{30(x^5 + 405x^4 + 27210x^3 + 488810x^2 + 2536005x + 3122577)}\\
		&+ \log{16}
	\end{split}
\end{align}

We note that the approximation in Equation~\ref{eq:optimal_log_approximation} requires the value of $\log{16}$. This can be computed in the plaintext domain and added to an encrypted value using the properties of the Paillier cryptosystem. The remaining arithmetic operations in the approximation can be computed in the encrypted domain, using the protocols described in Section~\ref{sec:fp_operations}. Therefore, this approximation can be calculated in a privacy-preserving manner.

Figure~\ref{fig:log_error_comparison} is a graph comparing the absolute error of the scaled approximation and the original approximation by Khattri given an input $x$. We can see that scaling the input by some constant factor $\alpha$ improves the accuracy of the closed-form approximation, although it requires that $\log \alpha$ be known.

\begin{figure}[ht]
	\centering
    \begin{tikzpicture}
		\begin{semilogyaxis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=300,
			ymin=1e-10, ymax=1.4,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		% \addplot[very thick, domain=0:300, samples=500]{abs(-1/30*(x^5*(30*ln(1/16) - 137) + 45*x^4*(270*ln(1/16) - 593) + 30*x^3*(27210*ln(1/16) - 20579) + 30*x^2*(488810*ln(1/16) + 216621) + 45*x*(1690670*ln(1/16) + 2694207) + 93677310*ln(1/16) + 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(x + 1))};
		% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
		\addplot[very thick] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
		\addplot[densely dashed, domain=0:300, samples=100]{abs(1/30*(137*x^5 + 2310*x^4 + 9870*x^3 + 15120*x^2 + 7560*x)/(x^5 + 30*x^4 + 210*x^3 + 560*x^2 + 630*x + 252) - ln(x + 1))};
		\legend{scaled ($\alpha = 1/16$), original ($\alpha = 1$)}
	\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Comparison of the absolute error between the original approximation in Equation~\ref{eq:standard_logarithm_quadrature} and the scaled approximation in Equation~\ref{eq:optimal_log_approximation}}
	\label{fig:log_error_comparison}
\end{figure}

We now show how we used numerical methods to determine that a scaling factor of $\alpha = 1/16$ minimizes the maximum absolute error of the scaled approximation for $\log{(1+x)}$, within the range $x \in [0, 255]$.
Let $I(x,\alpha)$ be the function obtained by approximating $\int_{n}^{\alpha n + \alpha}{\frac{1}{x}\diff x}$ with five-point Gauss--Legendre quadrature. The absolute error of the approximation compared to  $\log{(1+x)}$ at a given $x$ and $\alpha$ is given by the following function.
\begin{align*}
	A(x,\alpha) = \left| I(x,\alpha) - \log\alpha - \log{(1+x)}\right|.
\end{align*}

We want to show that the function
\begin{align*}
	f(\alpha) = \max_{0\leq x \leq 255}{A(x,\alpha)}
\end{align*}
is minimized at $\alpha = 1/16$.
We consider the functions $A(0,\alpha)$ and $A(255,\alpha)$. These are plotted in Figure~\ref{fig:endpoint_plot}.

It can be determined numerically using SageMath 8.3 that a local minimum of $\max(A(0,\alpha),A(255,\alpha))$ on the interval $[0,1]$ is at $\alpha = 1/16$, which is found at the intersection of the two graphs.
\begin{figure}[!ht]
	\centering
    \begin{tikzpicture}
		\begin{axis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=1,
			ymin=0, ymax=1.3,
			xlabel={$\alpha$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		\addplot[ACMDarkBlue, domain=0:1, samples=200]{1/30*abs((137*x^5 + 1625*x^4 + 2000*x^3 - 2000*x^2 - 30*(x^5 + 25*x^4 + 100*x^3 + 100*x^2 + 25*x + 1)*ln(x) - 1625*x - 137)/(x^5 + 25*x^4 + 100*x^3 + 100*x^2 + 25*x + 1))};
		\addplot[ACMOrange, domain=0:1, samples=200]{abs(-1/30*(32346117281250*x^5*ln(x) - 150633093005312*x^5 - 6979321856000*x^4 - 33554432000*x^3 + 131072000*x^2 + 30*(x^5 + 25*x^4 + 100*x^3 + 100*x^2 + 25*x + 1)*ln(x) + 38250*(x^5 + 20*x^4 + 60*x^3 + 40*x^2 + 5*x)*ln(x) + 19507500*(x^5 + 15*x^4 + 30*x^3 + 10*x^2)*ln(x) + 4974412500*(x^5 + 10*x^4 + 10*x^3)*ln(x) + 634237593750*(x^5 + 5*x^4)*ln(x) + 416000*x + 137)/(1099511627776*x^5 + 107374182400*x^4 + 1677721600*x^3 + 6553600*x^2 + 6400*x + 1) - 8*ln(2))};
		\legend{$x=0$, $x=255$}
	\end{axis}
	\end{tikzpicture}%
	\caption{Graphs of $A(x,\alpha)$ at $x=0$ and $x=255$}
	\label{fig:endpoint_plot}
\end{figure}
We then consider the function $A(x,1/16)$, shown in Figure~\ref{fig:single_alpha_plot}.
It was determined numerically using SageMath 8.3 that the local maxima of $A(x,1/16)$ on the interval occur at $x=0$ and $x=255$. Since a scaling factor other than $\alpha=1/16$ yields a higher absolute error at either $x=0$ or $x=255$, as seen in Figure~\ref{fig:endpoint_plot}, it follows that a scaling factor of $\alpha = 1/16$ minimizes the maximum absolute error of the scaled approximation for $\log{(1+x)}$, within the range $x \in [0, 255]$.
\begin{figure}[ht]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			xmin=0, xmax=300,
			ymin=0, ymax=0.035,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
		\addplot[very thick] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
	\end{axis}
	\end{tikzpicture}%
	\caption{Graph of $A(x,1/16)$, the absolute error of the approximation in Equation~\ref{eq:optimal_log_approximation} given a scaling factor of $\alpha=1/16$}
	\label{fig:single_alpha_plot}
\end{figure}

### 2019-01-18 pcsc-paper/parts/04_inversetrig.tex
Recreated graphs with pgfplots

\section{Inverse Trigonometric Functions}
We will now show how a Gauss-Legendre quadrature approach can also be used for efficient privacy-preserving approximation of the inverse tangent function. Two well-known series representations of the arctangent function exist. The Maclaurin series expansion,
\begin{align} \label{eq:arctan_maclaurin}
	\arctan x = \sum_{n=0}^{\infty}{\frac{(-1)^nx^{2n+1}}{2n+1}}, |x|\leq 1
\end{align}
and the following series expansion discovered by Euler in 1755, which converges more rapidly for all $x$ \cite{chien-lih_89.67_2005}.
\begin{align} \label{eq:arctan_euler}
	\arctan x = \sum_{n=0}^{\infty}
	{
	\frac
		{2^{2n}(n!)^2}
		{(2n+1)!}
	\frac
		{x^{2n+1}}
		{(1+x^2)^{n+1}}
	}
\end{align}
We will compare the quadrature approach with the method of computing partial sums from each of these series.

To obtain a Gauss-Legendre quadrature-based approximation, we take the following integral,
\begin{align*}
	\arctan x = \int_0^x{\frac{1}{1+x^2}\diff x},
\end{align*}
and perform five-point Gauss-Legendre quadrature to obtain the following closed-form approximation:
\begin{align} \label{eq:arctan_quadrature}
	\begin{split}
		&\arctan x \\
		&= \frac
		{4(225x^9 + 15925x^7 + 144753x^5 + 350595x^3 + 238140x)}
		{15(x^{10} + 480x^8 + 11760x^6 + 64120x^4 + 114660x^2 + 63504)}
	\end{split}
\end{align}

We note that in order to evaluate the quadrature approximation in equation \ref{eq:arctan_quadrature}, we must compute the value of $x^{10}$, which requires several rounds of the squaring and multiplication protocols to compute. We now compare the quadrature approximation to the partial sum composed of the first 4 terms of Euler's series expansion in equation \ref{eq:arctan_euler}, which requires a similar number of floating-point operations to compute:
\begin{align} \label{eq:arctan_euler_partial}
	\begin{split}
		&\sum_{n=0}^{3}
		{
		\frac
			{2^{2n}(n!)^2}
			{(2n+1)!}
		\frac
			{x^{2n+1}}
			{(1+x^2)^{n+1}}
		}\\
		&= \frac
		{965x^9 + 2370x^7 + 2688x^5 + 1470x^3 + 315x}
		{315(x^10 + 5x^8 + 10x^6 + 10x^4 + 5x^2 + 1)}
	\end{split}
\end{align}
Figure \ref{fig:arctan_error} shows that despite requiring a similar number of floating-point operations, the quadrature based approximation in equation \ref{eq:arctan_quadrature} gives accurate approximations over a larger interval, making it a more viable candidate for efficient computation.
\begin{figure}[!ht]
		\centering
		\includegraphics[width=.9\linewidth]{figures/single_alpha_plot.png}
		\caption{Comparison of the absolute error of the approximations for $\arctan x$ using quadrature approximation (equation \ref{eq:arctan_quadrature}) and the first 4 terms of Euler's series (equation \ref{eq:arctan_euler_partial})}
		\label{fig:arctan_error}
\end{figure}
Figure \ref{fig:arctan_error_2} shows the absolute error of the quadrature approximation over the interval $[-1,1]$. Since the quadrature approximation yields results with high accuracy over this interval, potential implmentations may use the following trigonometric identities to calculate values of the inverse tangent at larger arguments.
\begin{align*}
	\arctan x =
	\begin{cases}
		\frac{\pi}{2}-\arctan{\frac{1}{x}} & \text{if } x > 0, \\
		-\frac{\pi}{2}-\arctan{\frac{1}{x}} & \text{if } x < 0.
	\end{cases}
\end{align*}
To carry out the conditional check required to use these identities, a privacy-preserving comparison protocol may be used, such as in \cite{veugen_improving_2012}.
\begin{figure}[!ht]
		\centering
		\includegraphics[width=.9\linewidth]{figures/arctan_error_2.png}
		\caption{Absolute error of the approximation for $\arctan x$ using quadrature approximation (equation \ref{eq:arctan_quadrature}) over the interval $[-1,1]$}
		\label{fig:arctan_error_2}
\end{figure}

>>>>

\section{Inverse Trigonometric Functions}
We will now show how a Gauss--Legendre quadrature approach can also be used for efficient privacy-preserving approximation of the inverse tangent function. Two well-known series representations of the arctangent function exist: the Maclaurin series expansion,
\begin{align} \label{eq:arctan_maclaurin}
	\arctan x = \sum_{n=0}^{\infty}{\frac{(-1)^nx^{2n+1}}{2n+1}}, \text{ for }|x|\leq 1
\end{align}
and the following series expansion discovered by Euler in 1755, which converges more rapidly for all $x$~\cite{chien-lih_89.67_2005}.
\begin{align} \label{eq:arctan_euler}
	\arctan x = \sum_{n=0}^{\infty}
	{
	\frac
		{2^{2n}(n!)^2}
		{(2n+1)!}
	\frac
		{x^{2n+1}}
		{(1+x^2)^{n+1}}
	}
\end{align}
We will compare the quadrature approach with the method of computing partial sums from each of these series.

To obtain a Gauss--Legendre quadrature-based approximation, we take the following integral,
\begin{align*}
	\arctan x = \int_0^x{\frac{1}{1+x^2}\diff x},
\end{align*}
and perform five-point Gauss--Legendre quadrature to obtain the following closed-form approximation:
\begin{align} \label{eq:arctan_quadrature}
	\begin{split}
		&\arctan x \\
		&= \frac
		{4\left(225x^9 + 15925x^7 + 144753x^5 + 350595x^3 + 238140x\right)}
		{15\left(x^{10} + 480x^8 + 11760x^6 + 64120x^4 + 114660x^2 + 63504\right)}
	\end{split}
\end{align}

We note that in order to evaluate the quadrature approximation in Equation~\ref{eq:arctan_quadrature}, we must compute the value of $x^{10}$, which requires several rounds of the squaring and multiplication protocols to compute. We now compare the quadrature approximation to the partial sum composed of the first 4 terms of Euler's series expansion in Equation~\ref{eq:arctan_euler}, which requires a similar number of floating-point operations to compute:
\begin{align} \label{eq:arctan_euler_partial}
	\begin{split}
		&\sum_{n=0}^{3}
		{
		\frac
			{2^{2n}(n!)^2}
			{(2n+1)!}
		\frac
			{x^{2n+1}}
			{(1+x^2)^{n+1}}
		}\\
		&= \frac
		{965x^9 + 2370x^7 + 2688x^5 + 1470x^3 + 315x}
		{315\left(x^{10} + 5x^8 + 10x^6 + 10x^4 + 5x^2 + 1\right)}
	\end{split}
\end{align}

Figure~\ref{fig:arctan_error} shows that despite requiring a similar number of floating-point operations, the quadrature based approximation in Equation~\ref{eq:arctan_quadrature} gives accurate approximations over a larger interval, making it a more viable candidate for efficient computation.
\begin{figure}[!ht]
		\centering
		\includegraphics[width=.9\linewidth]{figures/single_alpha_plot.png}
		\caption{Comparison of the absolute error of the approximations for $\arctan x$ using quadrature approximation (Equation~\ref{eq:arctan_quadrature}) and the first 4 terms of Euler's series (Equation~\ref{eq:arctan_euler_partial})}
		\label{fig:arctan_error}
\end{figure}

Figure~\ref{fig:arctan_error_2} shows the absolute error of the quadrature approximation over the interval $[-1,1]$. Since the quadrature approximation yields results with high accuracy over this interval, potential implmentations may use the following trigonometric identities to calculate values of the inverse tangent at larger arguments.
\begin{align*}
	\arctan x =
	\begin{cases}
		\frac{\pi}{2}-\arctan{\frac{1}{x}} & \text{if } x > 0 \\
		-\frac{\pi}{2}-\arctan{\frac{1}{x}} & \text{if } x < 0 
	\end{cases}
\end{align*}

To carry out the conditional check required to use these identities, a privacy-preserving comparison protocol may be used, such as in~\cite{veugen_improving_2012}.
\begin{figure}[!ht]
		\centering
		\includegraphics[width=.9\linewidth]{figures/arctan_error_2.png}
		\caption{Absolute error of the approximation for $\arctan x$ using quadrature approximation (Equation~\ref{eq:arctan_quadrature}) over the interval $[-1,1]$}
		\label{fig:arctan_error_2}
\end{figure}
